### 2019-06-07 reports/thesis/chapters/2_dataCollection.tex
chapter 2 writing

The shading free image $U_{x,y}$ is not truly known, but it should expected that the shading corrected greyscale projection should be flat with some noise, in other words $\expectation\left[S_{x,y}(P)\right]=\mu_{S}(P)$ and $\variance\left[S_{x,y}(P)\right]=\sigma_{S}^2(P)$ for all $x=1,2,\cdots,L$ and $y=1,2,\cdots,H$. Suppose $\mu_{S}(P)$ was estimated using the within projection mean
\begin{equation}
\overline{S}(P) = \dfrac{1}{\text{WH}}
\sum_{x=1}^W\sum_{y=1}^H S_{x,y}(P) \ .
\end{equation}
Consider a pixel at $(x,y)$, shading correction was done by fitting a linear regression on
\begin{equation}
\overline{S}(P) = \beta_{x,y} S_{x,y}(P) + \alpha_{x,y} \quad(+\varepsilon)
\end{equation}
for $P\in \mathfrak{P}$ where $\mathfrak{P} = \left\{0,P_1,P_2,\cdots, P_\text{proj}\right\}$ are the powers used for the greyscale projections. $\varepsilon$ is a random variable and an error term, it is included only for formality purposes. Let $b_{x,y}$ and $a_{x,y}$ be the estimated parameters of $\beta_{x,y}$ and $\alpha_{x,y}$ from the linear regression respectively. Given a projection $N_{x,y}$, the shading corrected projection $U_{x,y}$ is
\begin{equation}
U_{x,y} = b_{x,y} N_{x,y} + a_{x,y}
\end{equation}
or in the form of Equation \eqref{eq:data_shadingCorrectionOld}
\begin{multline}
U_{x,y} = \dfrac{N_{x,y}-\text{black}_{x,y}}{\text{white}_{x,y}-\text{black}_{x,y}}\times \left[b_{x,y}(\text{white}_{x,y}-\text{black}_{x,y})\right]\\+\left[b_{x,y} \text{black}_{x,y} + a_{x,y}\right] \ .
\end{multline}
In full, the equations for $b_{x,y}$ and $a_{x,y}$ are given as
\begin{equation}
b_{x,y} = \dfrac{
  \sum_{P\in\mathfrak{P}}(S_{x,y}(P) - \overline{S}_{x,y})(\overline{S}(P) - \overline{S})
}{
  \sum_{P\in\mathfrak{P}}(S_{x,y}(P) - \overline{S}_{x,y})^2
}
\end{equation}
and
\begin{equation}
a_{x,y} = \overline{S} - b_{x,y}\overline{S}_{x,y}
\end{equation}
where $\overline{S}_{x,y}$ is the between projection mean
\begin{equation}
\overline{S}_{x,y} = \dfrac{1}{|\mathfrak{P}|}\sum_{P\in\mathfrak{P}}S_{x,y}(P)
\end{equation}
and $\overline{S}$ is the global mean
\begin{equation}
\overline{S} = \dfrac{1}{|\mathfrak{P}|}\sum_{P\in\mathfrak{P}}\overline{S}(P) \ .
\end{equation}
This type of shading correction shall be referred to as linear shading correction. Expressing shading correction in this way has the advantage that there are no user defined constants.

An example of the linear regression is shown in Figure \ref{fig:data_shadingCorrectionExample_gainMap} where 3 random pixels were chosen for illustration. By plotting the within projection mean versus the grey value for a particular pixel, a linear relationship can be observed. The gradient varied for different pixels which correspond to different sensitivities. The resulting shading correction for the \texttt{AbsNoFilter} projection is shown in Figure \ref{fig:data_shadingCorrectionExample_image}. It can be observed that the shading correction remove panel and spot effects from the background and test object.

@@

\begin{figure}
  \centering
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_image_null.eps}
      \caption{No shading correction}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_image_bw.eps}
      \caption{Linear shading correction}
    \end{subfigure}
  }
  \caption{Projection of \texttt{AbsNoFilter} with and without shading correction}
  \label{fig:data_shadingCorrectionExample_image}
\end{figure}

@@

@@

\begin{figure}
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_power_null.eps}
      \caption{No shading correction}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_power_bw.eps}
      \caption{bw shading correction}
    \end{subfigure}
  }
  \caption{Gain curve before and after shading correction. The box plot summarise all $2000\times2000$ pixels in a projection.}
  \label{fig:data_shadingCorrectionExample_gain}
\end{figure}

@@

\begin{figure}
  \centerline{
    \begin{subfigure}[b]{\textwidth}
      \includegraphics[width=0.49\textwidth]{../figures/data/oddEvenPlot1_null.eps}
      \includegraphics[width=0.49\textwidth]{../figures/data/oddEvenPlot2_null.eps}
      \caption{No shading correction}
    \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{\textwidth}
      \includegraphics[width=0.49\textwidth]{../figures/data/oddEvenPlot1_bw.eps}
      \includegraphics[width=0.49\textwidth]{../figures/data/oddEvenPlot2_bw.eps}
      \caption{bw shading correction}
    \end{subfigure}
  }
  \caption{Left shows the profile plot of a black image at $(879,y)$ before and after shading correction. Right shows two curves for odd and even $y$.}
  \label{fig:data_oddEven}
\end{figure}

\begin{figure}
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_blackFft_null.eps}
      \caption{No shading correction}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_blackFft_bw.eps}
      \caption{bw shading correction}
    \end{subfigure}
  }
  \caption{Fourier transform on the black image before and after shading correction. The colour scale is in arbitrary units and range between 0 to 255.}
  \label{fig:data_fft}
\end{figure}

\begin{figure}
  \centerline{
    \begin{subfigure}[b]{\textwidth}
      \includegraphics[width=0.49\textwidth]{../figures/data/oddEvenPlot1_nullX.eps}
      \includegraphics[width=0.49\textwidth]{../figures/data/oddEvenPlot2_nullX.eps}
      \caption{No shading correction}
    \end{subfigure}
  }
  \caption{Left shows the profile plot of a black image at $(y,879)$ before and after shading correction. Right shows two curves for odd and even $y$.}
  \label{fig:data_oddEvenX}
\end{figure}

>>>>

The shading free image $U_{x,y}$ is not truly known, but it should expected that the shading corrected greyscale projection should be flat with some noise, in other words $\expectation\left[S_{x,y}(P)\right]=\mu_{S}(P)$ and $\variance\left[S_{x,y}(P)\right]=\sigma_{S}^2(P)$ for all $x=1,2,\cdots,L$ and $y=1,2,\cdots,H$. Suppose $\mu_{S}(P)$ was estimated using the within projection mean
\begin{equation}
\overline{S}(P) = \dfrac{1}{\text{WH}}
\sum_{x=1}^W\sum_{y=1}^H S_{x,y}(P) \ .
\end{equation}
Consider a pixel at $(x,y)$, shading correction was done by fitting a linear regression on
\begin{equation}
\overline{S}(P) = \beta_{x,y} S_{x,y}(P) + \alpha_{x,y} \quad(+\varepsilon)
\end{equation}
for $P\in \mathfrak{P}$ where $\mathfrak{P} = \left\{0,P_1,P_2,\cdots, P_\text{proj}\right\}$ are the powers used for the greyscale projections. $\varepsilon$ is a random variable and an error term, it is included only for formality purposes. Let $b_{x,y}$ and $a_{x,y}$ be the estimated parameters of $\beta_{x,y}$ and $\alpha_{x,y}$ from the linear regression respectively. Given a projection $N_{x,y}$, the shading corrected projection $U_{x,y}$ is
\begin{equation}
U_{x,y} = b_{x,y} N_{x,y} + a_{x,y} \ .
\end{equation}
In full, the equations for $b_{x,y}$ and $a_{x,y}$ are given as
\begin{equation}
b_{x,y} = \dfrac{
  \sum_{P\in\mathfrak{P}}(S_{x,y}(P) - \overline{S}_{x,y})(\overline{S}(P) - \overline{S})
}{
  \sum_{P\in\mathfrak{P}}(S_{x,y}(P) - \overline{S}_{x,y})^2
}
\end{equation}
and
\begin{equation}
a_{x,y} = \overline{S} - b_{x,y}\overline{S}_{x,y}
\end{equation}
where $\overline{S}_{x,y}$ is the between projection mean
\begin{equation}
\overline{S}_{x,y} = \dfrac{1}{|\mathfrak{P}|}\sum_{P\in\mathfrak{P}}S_{x,y}(P)
\end{equation}
and $\overline{S}$ is the global mean
\begin{equation}
\overline{S} = \dfrac{1}{|\mathfrak{P}|}\sum_{P\in\mathfrak{P}}\overline{S}(P) \ .
\end{equation}
This type of shading correction shall be referred to as linear shading correction. Expressing shading correction in this way has the advantage that there are no user defined constants.

An example of the linear regression is shown in Figure \ref{fig:data_shadingCorrectionExample_gainMap} where 3 random pixels were chosen for illustration. By plotting the within projection mean versus the grey value for a particular pixel, a linear relationship can be observed. The gradient varied for different pixels which correspond to different sensitivities. The resulting shading correction for the \texttt{AbsNoFilter} projection is shown in Figure \ref{fig:data_shadingCorrectionExample_image}. It can be observed that the shading correction removed the panel and spot effects from the background and test object.

@@

\begin{figure}
  \centering
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_image_null.eps}
      \caption{No shading correction}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_image_linear.eps}
      \caption{Linear shading correction}
    \end{subfigure}
  }
  \caption{Projection of \texttt{AbsNoFilter} with and without shading correction}
  \label{fig:data_shadingCorrectionExample_image}
\end{figure}

@@

\subsection{Exploratory Analysis}

Shading correction on a greyscale projection should result in a image which is flat with some noise. To avoid overfitting, one greyscale projection from each power were held out and used to fit the parameters of the shading correction. Shading correction was then used on unused greyscale projections in this exploratory analysis.

Figure \ref{fig:data_shadingCorrectionExample_greyvaluePower} shows the grey values in each greyscale projections before and after shading correction. The sensitivity, which corresponds to the gradient in units of \SI{}{\adu\per\watt}, became more consistent between pixels with shading correction. This should mean that pixels will respond similarly to each other.

\begin{figure}
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_power_null.eps}
      \caption{No shading correction}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_power_linear.eps}
      \caption{Linear shading correction}
    \end{subfigure}
  }
  \caption{Grey values in the \texttt{AbsNoFilter} greyscale projections before and after shading correction. The box plot summarise all $2000\times2000$ pixels in a projection.}
  \label{fig:data_shadingCorrectionExample_greyvaluePower}
\end{figure}

The \texttt{AbsNoFilter} black image before and after shading correction is shown in Figure \ref{fig:data_shadingCorrectionExample_black}. The flatness of the image can be shown using a profile plot, this is a plot of the grey values along a column or row, an example shown in Figure \ref{fig:data_oddEven}. The figure shows that the shading uncorrected black image is not flat but a remarkable structure can be observed by plotting only the odd rows and even rows separately. Such a plot shows that the grey values depend on neighbouring pixels, where the majority of pixels on even rows have grey values larger than pixels above and below it. Perhaps this is caused by the read-out structure in the detector.

The BW shading corrected black image looks very uniform but there is some structure in the linear shading correction. By giving linear shading correction various greyscale projections, it tried to generalise to a range of powers and thus may struggle at shading correcting the black and white images.

\begin{figure}
  \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_black_bw.eps}
      \caption{BW shading correction}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_black_linear.eps}
      \caption{Linear shading correction}
    \end{subfigure}
  }
  \caption{The \texttt{AbsNoFilter} black image after shading correction.}
  \label{fig:data_shadingCorrectionExample_black}
\end{figure}

\begin{figure}
  \centerline{
    \begin{subfigure}[b]{\textwidth}
      \includegraphics[width=0.49\textwidth]{../figures/data/oddEvenPlot1_null.eps}
      \includegraphics[width=0.49\textwidth]{../figures/data/oddEvenPlot2_null.eps}
      \caption{No shading correction}
    \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{\textwidth}
      \includegraphics[width=0.49\textwidth]{../figures/data/oddEvenPlot1_bw.eps}
      \includegraphics[width=0.49\textwidth]{../figures/data/oddEvenPlot2_bw.eps}
      \caption{BW shading correction}
    \end{subfigure}
  }
  \centerline{
    \begin{subfigure}[b]{\textwidth}
      \includegraphics[width=0.49\textwidth]{../figures/data/oddEvenPlot1_linear.eps}
      \includegraphics[width=0.49\textwidth]{../figures/data/oddEvenPlot2_linear.eps}
      \caption{linear shading correction}
    \end{subfigure}
  }
  \caption{Left shows the profile plot of a \texttt{AbsNoFilter} black image at $(879,y)$. Right shows two curves for odd and even $y$.}
  \label{fig:data_oddEven}
\end{figure}

% \begin{figure}
%   \centering
%   \centerline{
%     \begin{subfigure}[b]{0.49\textwidth}
%       \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_blackFft_null.eps}
%       \caption{No shading correction}
%     \end{subfigure}
%     \begin{subfigure}[b]{0.49\textwidth}
%       \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_blackFft_bw.eps}
%       \caption{BW shading correction}
%     \end{subfigure}
%   }
%   \begin{subfigure}[b]{0.49\textwidth}
%     \includegraphics[width=\textwidth]{../figures/data/shadingCorrectionExample_blackFft_linear.eps}
%     \caption{Linear shading correction}
%   \end{subfigure}
%   \caption{Fourier transform on the black image before and after shading correction. The colour scale is in arbitrary units and range between 0 to 255.}
%   \label{fig:data_fft}
% \end{figure}

@@

@@
